---
output: github_document
bibliography: vignettes/References.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/MWTools)](https://CRAN.R-project.org/package=MWTools)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Codecov test coverage](https://codecov.io/gh/EnergyEconomyDecoupling/MWTools/branch/master/graph/badge.svg)](https://codecov.io/gh/EnergyEconomyDecoupling/MWTools?branch=master)
[![R-CMD-check](https://github.com/EnergyEconomyDecoupling/MWTools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/EnergyEconomyDecoupling/MWTools/actions/workflows/R-CMD-check.yaml)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.7584859.svg)](https://doi.org/10.5281/zenodo.7584859)
<!-- badges: end -->


# MWTools

## Statement of need

In societal energy analysis and societal exergy analysis (SEA), 
animals and humans provide energy to society by performing _muscle work_.
A previous study by [@Steenwyk:2022ww](https://doi.org/10.1007/s41247-022-00096-z)
developed and standardized ways to estimate
muscle work contributed to society by animals and humans.
And data exist from the UN's
[Food and Agriculture Organization](https://www.fao.org/home/en) (FAO) and
the [International Labor Organization](https://www.ilo.org) (ILO)
that could be used for such purposes.
However, none of the data are in the PSUT format
proposed by 
[@Heun:2018](https://doi.org/10.1016/j.apenergy.2018.05.109).
Computational tools are needed to gather and shape relevant data.

This package provides functions for the estimation 
of animal and human muscle work for use in SEA
with the Physical Supply Use Table (PSUT) framework.


## Installation

You can install this package from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("EnergyEconomyDecoupling/MWTools")
```


## More Information

Find more information, including vignettes and function documentation, at 
<https://energyeconomydecoupling.github.io/MWTools/>.


## References
