---
title: "MWTools"
output: rmarkdown::html_vignette
bibliography: References.bib
vignette: >
  %\VignetteIndexEntry{MWTools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

In societal energy analysis and societal exergy analysis (SEA), 
animals and humans provide energy to society by performing _muscle work_.
A previous study by [@Steenwyk:2022ww](https://doi.org/10.1007/s41247-022-00096-z)
developed and standardized ways to estimate
muscle work contributed to society by animals and humans.
And data exist from the UN's
[Food and Agriculture Organization](https://www.fao.org/home/en) (FAO) and
the [International Labor Organization](https://www.ilo.org) (ILO)
that could be used for such purposes.
However, none of the data are in the PSUT format
proposed by 
[@Heun:2018](https://doi.org/10.1016/j.apenergy.2018.05.109).
Computational tools are needed to gather and shape relevant data.

The examples below show how to use `MWTools` to address
the needs described above.

## Examples

### Animal Muscle Work

Raw data for the estimation of animal muscle work is obtained from the Food and 
Agriculture Organisation of the United Nations Statistical Database (FAOSTAT), 
via the `R` package `FAOSTAT` as follows.

```{r down_amw_example, eval = FALSE}
raw_amw_data <- FAOSTAT::get_faostat_bulk(code = "QCL",  data_folder = tempdir())
```

The helper function `calc_amw_pfu` returns a tidy data frame containing data for
the primary, final, and useful energy by species, country, year and in
the agriculture, transport, and total sectors.

```{r calc_amw_example}
test_amw_data <- read.csv(file = MWTools::amw_test_data_path())

amw_pfu_data <- MWTools::calc_amw_pfu(.df = test_amw_data,
                                      concordance_path = MWTools::fao_concordance_path(),
                                      amw_analysis_data_path = MWTools::amw_analysis_data_path())
tibble::glimpse(amw_pfu_data)
```

### Human Muscle Work

Raw data for the estimation of human muscle work is obtained from the 
International Labor Organisation (ILO), via the `R` package `Rilostat` as 
follows: 

```{r down_hmw_example, eval = FALSE}
ilo_working_hours_data <- Rilostat::get_ilostat(id = MWTools::ilo_codes$working_hours_code,
                                                quiet = TRUE) |>
  Rilostat::label_ilostat(code = c("ref_area"))

ilo_employment_data <- Rilostat::get_ilostat(id = MWTools::ilo_codes$employment_code,
                                             quiet = TRUE) |>
  Rilostat::label_ilostat(code = c("ref_area"))
```

Using data for the number of employed persons and mean yearly working hours the 
primary, final, and useful energy associated with human muscle work can be 
estimated using the helper function `MWTools::calc_hmw_pfu` as below.

``` {r calc_hmw_example}
ilo_working_hours_data <- read.csv(file = MWTools::ilo_working_hours_test_data_path())
ilo_employment_data <- read.csv(file = MWTools::ilo_employment_test_data_path())

preparedILOData <- MWTools::prepareRawILOData(ilo_working_hours_data = ilo_working_hours_data,
                                              ilo_employment_data = ilo_employment_data)

hmw_pfu_data <- MWTools::calc_hmw_pfu(.df = preparedILOData,
                                      concordance_path = MWTools::fao_concordance_path(),
                                      hmw_analysis_data_path = MWTools::hmw_analysis_data_path())
tibble::glimpse(hmw_pfu_data)
```


### PSUT matrices

PFU data can be converted to PSUT matrices using `prep_psut()`.

```{r prep_psut_example}
mw_psut <- MWTools::prep_psut(.hmw_df = hmw_pfu_data,
                              .amw_df = amw_pfu_data)
tibble::glimpse(mw_psut)
```


## Conclusion

The functions in the `MWTools` package can streamline and assist inclusion 
of animal and human muscle work in societal exergy analysis.
