---
title: "MWTools"
output: rmarkdown::html_vignette
bibliography: References.bib
vignette: >
  %\VignetteIndexEntry{IEATools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

In societal energy analysis and societal exergy analysis (SEA), 
animals and humans provide energy to society by performing _muscle work_.
A previous study by [@Steenwyk:2022ww](https://doi.org/10.1007/s41247-022-00096-z)
developed and standardized ways to estimate
muscle work contributed to society by animals and humans.
And data exist from the UN's
[Food and Agriculture Organization](https://www.fao.org/home/en) (FAO) and
the [International Labor Organization](https://www.ilo.org) (ILO)
that could be used for such purposes.
However, none of the data are in the PSUT format
proposed by 
[@Heun:2018](https://doi.org/10.1016/j.apenergy.2018.05.109).
Computational tools are needed to gather and shape relevant data.

The examples below show how to use `MWTools` to address
the needs described above.


## Examples

### Animal Muscle Work

Raw data for the estimation of animal muscle work
is obtained from the Food and Agriculture Organisation 
of the United Nations Statistical Database (FAOSTAT), 
via the `R` package `FAOSTAT`.
The `MWTools` package provides a wrapper function 
for downloading data for the number of live animals,
`down_fao_live_animals()`, which utilises the function `FAOSTAT::get_faostat_bulk()`. 
Once downloaded, `MWTools` provides a number of functions
for the calculation of the number of working animals 
by species, country, year and in agriculture, transport, and in total.

The helper function `calc_amw_pfu()` returns a tidy data frame
containing data for the primary, final, and useful energy 
by species, country, year and in agriculture, transport, and in total.

```{r get_example_data}
library(MWTools)

fao_fp <- PFUSetup::get_abs_paths(version = "v1.2")$fao_data_path

fao_raw_rds <- readr::read_rds(file = fao_fp)

amw_pfu_data <- calc_amw_pfu(.df = fao_raw_rds,
                             concordance_path = PFUSetup::get_abs_paths()$mw_concordance_path,
                             amw_analysis_data_path = PFUSetup::get_abs_paths()$amw_analysis_data_path)

head(amw_pfu_data)
```


### Human Muscle Work

Raw data for the estimation of human muscle work
is obtained from the International Labor Organisation (ILO), 
via the `R` package `Rilostat`.
The `MWTools` package includes bundled ILO data for the number of employed persons by sector and mean number of working hours by sector.

Using data for the number of employed persons and 
mean yearly working hours,
the primary, final, and useful energy associated 
with human muscle work can be estimated 
using the helper function `calc_hmw_pfu()`.

``` {r hmw_pfu}
library(MWTools)

ilo_working_hours_data <- readr::read_rds(file = PFUSetup::get_abs_paths(version = "v1.2")$ilo_working_hours_data_path)
ilo_employment_data <- readr::read_rds(file = PFUSetup::get_abs_paths(version = "v1.2")$ilo_employment_data_path)

preparedILOData <- prepareRawILOData(ilo_working_hours_data = ilo_working_hours_data,
                                     ilo_employment_data = ilo_employment_data)

hmw_pfu_data <- calc_hmw_pfu(.df = preparedILOData,
                             concordance_path = PFUSetup::get_abs_paths()$mw_concordance_path,
                             hmw_analysis_data_path = PFUSetup::get_abs_paths()$hmw_analysis_data_path)

head(hmw_pfu_data)
```


### PSUT matrices

PFU data can be converted to PSUT matrices using `prep_psut()`.


## Conclusion

The functions in the `MWTools` package can streamline and assist inclusion 
of animal and human muscle work in societal exergy analysis.
